\documentclass[a4paper, 12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{ebgaramond}
\usepackage{amsmath, amssymb}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example} \newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}


\begin{document}

The night falls deep: I sense the moonlight with my heart. Again, I cannot
sleep. What troubles me are these pearls, these little devils:
\textit{memories}. Though taking many forms, like ill-intended water, a face, a
name, a man always appears. His eyes are a thousand eyes, his face is a
thousand faces, his soul is a thousand souls. Violence and love, tranquility
and vengeance, obscurity and silence, weariness and hope: these are his
avatars. 

These which I here collect are fragments of a man as seen through the eyes of
his youngest son. Whatever I may say cannot do justice neither to his clarities
nor his obscurities. So I will refrain myself from commenting on these memories
as they appear to me \textit{today}, and will attempt to simply write them down
as they are in themselves.

~ 

$\S$ L and I are playing and swimming in the waters. L lost his father when he
was four. At sunset we head back to my house: Again, as it was usual in the
hellish summers of Corrientes, we have no electricity. A familiar dilemma is
faced: Open the windows and allow the nightly breeze to flow, but with it the
innumerable mosquitoes; keep the windows closed, the plague at bay,
and bare the heat. My father calls us in from his room: In the little terrace,
three mattresses are lined up next to one another. Above them, because we are 
far from the city and there are no lights in the blackout, a myriad of stars of every
color blaze. To our surprise, the breeze is strong enough to keep the mosquitoes 
away. As we all lie there, in the terrace, without blankets or pillows,
my father begins to speak: \textit{Those are the} Tres Marías, \textit{that is } Orión...
And so we fall asleep.

~ 


$\S$ My sister and I are lying in bed with mom and dad. We play a game called
\textit{syllables}: My father will speak a word, we will split it into
syllables, and when the syllables form a bad word he will tickle us. The room
is dark but I can hear my mother laugh and feel the silk of her clothes. My
sister and I tremble when a word contains a bad word, as we know the faith
awaiting us.

~ 

$\S$ My father is driving my siblings and I around the city. I'm maybe nine
years old. He explains that he's divorcing my mother. I cannot recollect
exactly what happened in-between, but now I have his phone in my hand and I am
calling my mother. When I hear her voice I start to cry, and when she asks
what's the matter, I can only mumble: \textit{Es que vos y papá...}. And I hear
that she cries too.

~ 

$\S$ My siblings, my grandmother and I are in the car with my father. She
already shows the symptoms of dementia. We drive her to her nursing home. My
father walks her in as we wait in the car. When he gets back he puts his arms
on the wheel and sinks his face in them. It is the first time that I hear
him cry. He says: \textit{Perdón, chicos, es que su mamá... y mi mamá...} My
sister and I, who are in the back seats, rush to comfort him; we pad and caress
his back and say that everything will be fine. My brother, in the front seat,
rolls his eyes and bites his lips, displaying anger or frustration.

~

$\S$ It is Sunday morning, which means my father will pick us up. As usual, my
brother and sister are either asleep or pretending to be asleep so as not to
go. (At the time, I judged them: Now I see that they lacked, for better or for
worse, the capacity to ignore the bitter remarks my father always made
about my mother.) Now I am with my father in \textit{Puente Pexoa}, our usual
fishing spot. He is an good mood, but at times he becomes somber and quiet, and
appears to be angry. I almost catch a large \textit{Yarú Itacuá}
(\textit{Plecostomus commersonii}) but the beast is able to free itself. We
see, in a small islet, a group of men celebrating: They caught a
\textit{yacaré} and will have it for dinner. The small alligator seems to have
accepted his faith.

~ 

$\S$ A few weeks ago I hurt one of my toes playing with T. I have complained
repeatedly to my mother but she keeps telling me it will pass. It's late at
night, after midnight, and a lot of people are in my home---I don't know who
they are---. I'm sitting in what used to be my father's studio and I feel
complete desolation. I call my father and ask him to come pick me up. After
perhaps twenty minutes he arrives at the corner. I pick up
my guitar---I had began my studies already---and limp to his car. He asks why
is it that I'm limping and I tell him of the accident. He shows frustration
that my mother didn't take me to the doctor. I explain to him my
feelings---these were complex and real, though I was only twelve years old--- I
speak of desolation, of anguish, and despair. He understands what I'm
suggesting and makes the first move, asking if I've considered moving in with
him. I told him that was what I wanted. And so the most divisive and important
decision in my life was made---one whose repercussions I never even began to
fathom.

~ 

$\S$ It is the day after to the preceding memory. I ask my father to notify my
mother of my decision to live with him, under the condition that he does so in
front of me. When he tells my mother that I'm going to live with him from then
on, she simply asks how will the arrangements be for picking me up from school.
(The coldness of my mother's reply haunted me for years and, at the moment, I
took them as a confirmation of the indifference with which she seemed to live
the crisis. Many years later, I learned that after that call she spent days
without leaving her bed and without ceasing to cry.)

~ 

$\S$ I have started to show an inclination for poetry. I'm writing and reading
everyday. My father knocks on my door and comes with an immense tower of books.
Garcilaso, Lezama Lima, Borges, Shelley... He's gifting me all his poetry
books.

~ 

$\S$ I have a falling-out with my mother: I am angry and confused. (These are
the constant epithets of all my adolescence.) I speak to my father obscure and
foolish words. He's quiet, sitting, statuary and grim: and I see that he is
rejoicing.

~ 

$\S$ The day after a significant fight, in which all family members were
involved, I wake up. I go to my father's room and the first thing he does is
scream to me that my mother is a whore.





































\end{document}



