\documentclass[a4paper, 12pt]{article}
\usepackage{verse, gmverse}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{newtxtext} \usepackage{newtxmath}
\usepackage{titlesec}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{}{1em}{}
\titleformat{\subsection}[hang]{\normalfont\filcenter\MakeUppercase}{}{1em}{}
\setcounter{secnumdepth}{0}


\begin{document}

\subsection{I}

   there is a truth we cannot bear

   oh, all the drowning men have seen it

   their foaming eyes have almost touched it 

   in all their hearts it was a ruby 

   it was a moon as red as sunset's dying zest

   ~ ~ ~ 

   if you have walked, a shadow of your shadow,
   
   a ghost quadruped, primitive---not feral, but still close---

   if you have listened to the beggar's begging

   when sitting in your bed at night, alone--- 

   if you have heard a whisper cry 
   
   a piece of bread, for love of god...!

   and turned to see the darkness in your room---

   ~ ~ ~ 

   if you have loved I mean, a love outcasted, caged,

   then yes, you know the truth we cannot bear 

   you know there is a sea of drowning men 

   and in your heart there is a ruby 

   a violent moon you can't give up

   however much its weight may search the kiss 

   of sandless abysses of time




   \pagebreak

   \subsection{II}

How longingly I have wished to find you here in the solitary corridors of dream.
Do you recall the secret which was my offering to you? Do you recall how young
the dreams I dreamt, how juvenile the shyness of my eyes? Remember now how well
it stood the truth there dripping from my lips. Now I have learnt that love is
ineternal---for everyone, and for us too---. The flow of life is irresistible.
Nothing is true but for a day.

I wrote to you one day that I was like the wind. Draw your curtains open, I
said, and I shall flow to you, so that you may become a stranger to solitude. 
We were but children. The moon was shining clear. Your face was that of early
youth. And now I cannot draw my mind away from thinking about the windlessness of
the earth.

To love is to uphold the entirety of a being, as if it was a final offering, the
one and only sacrifice which could bargain the forgiveness of our god. Here in
my house, alone, I sit alone and write at night, alone, that I uphold you.
Still, after all these years... I speak your name so that you become a spell. I
vividly imagine you, so that you become a ghost. 

I must remember you, so as to live this searching life which gives the memory of
you.

\pagebreak 

\begin{verse}
    
mi alma nunca está desnuda\\

\\ 

mi alma es como un sátiro nocturno
que odia la luz y siente que la luna es superflua

es una dura piedra milenaria
que en su interior esconde un agua negra

un árbol para quien la savia oscura
es el total secreto de su vida

pocos la ven, yo mismo no la he visto,
pero la intuyo en turbias confesiones
que no sé pronunciar pero confieso




\end{verse}

\pagebreak

\begin{verse}
    He olvidado mi lengua, mi voz y mi lenguaje.
    Mi nombre silenciado no es dicho: es intuido.
    Sólo habla entre nosotros el místico oleaje 
    de un signo primitivo, de un idioma extinguido.
    (Esta adicción enferma no es odio: es lo que traje
    del mundo analfabeto y sin habla de tu olvido.)
    ¿Recuerdas aquel sueño? "Un pájaro salvaje 
    le obsequia a la serpiente su corazón vencido,
    y desde entonces vive afónico y desnudo".
    (El pájaro del sueño ya está yaciendo frío,
    presa de la serpiente, del hongo y el gusano.)
    Y sin embargo adoro este universo mudo:
    por lo menos es algo nuestro, tuyo y mío,
    y es mi mordaza el turbio fantasma de tu mano.
    \end{verse}

\pagebreak


\pagebreak

\begin{verse}
    He olvidado mi lengua, mi voz y mi lenguaje.
    Mi nombre silenciado no es dicho: es intuido.
    Sólo habla entre nosotros el místico oleaje 
    de un signo primitivo, de un idioma extinguido.

    Esta adicción enferma no es odio: es lo que traje
    del mundo analfabeto y sin habla de tu olvido.
    (Allí soy como el sueño de un pájaro salvaje 
    que obsequia a la serpiente su corazón vencido.)

    Sigamos en silencio, tu allá, yo aquí, distante.
    (El pájaro del sueño ya está yaciendo frío,
    presa de la serpiente, del hongo y el gusano.)

    Yo soy feliz en este silencio de diamante:
    por lo menos es algo nuestro, tuyo y mío,
    y es mi mordaza el turbio fantasma de tu mano.

    \end{verse}

\pagebreak

\begin{verse}
   No digas nada: olvida tu lengua y tu lenguaje.
   Si algo puede ser dicho, puede ser intuido. $\qquad$ (?)
   Sintamos el idioma del místico oleaje 
   que fluye entre nosotros cual lenguaje extinguido.

   Sólo esqueletos quedan de todo aquel plumaje 
   que revistió el encuentro de mi alma con tu olvido.
   Y aquellas plumas negras, y aquel nocturno traje
   que disfrazó mi vida, odiaban el sonido.

   Sigamos en silencio, tú allá, yo aquí, distante.
   Mis penas inauditas se han ido con el viento 
   y las recuerdo apenas como un rostro lejano.

   Sólo recuerdo ahora un silencio de diamante
   que honro sellando a piedra el pulso de mi aliento.
   Y es mi mordaza el turbio fantasma de tu mano.
   

\end{verse}

\pagebreak 

\begin{verse}
    here lies the moon trapped in the fist of men 
    here yearns the sun for freedom from the night 
    here in the flesh once known to you the sinister is shackled
    here in these eyes that orbit 'round your grave these tears

    these are the things I keep
    the things I captive took a morning of rapacious rage
    the harvest of the earth that chambers you

    these are the things with which I am awaiting 
    these are the evil tokens of my anticipation
    with these I wait the day of our reunion 
    in timeless spaceless unicorn obscurity 


\end{verse}




\pagebreak 


\begin{verse}
    
La lluvia melodiosa forma charcos 
de culpa en mi silencio. ¿Me conoces
o te olvidaste de estos ojos parcos?
~
Todo está lleno de almas y de dioses
en este jardín plástico y espurio
que ofende la tibieza de mi vida.
~
¿Tienes miedo? ¿Te asusta un mal augurio?
¿Sientes el frío de la lluvia herida?
No: todo lo ignoras. Nada es bello
ni atroz ante tus manos. Este frío 
hiere unos huesos que has desocupado.
~
El nudo que agobió una vez tu cuello
quedó cerrado siempre sobre el mío.
~
(Todo es gradual y eterno de este lado.)
\end{verse}


















\end{document}



