\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
\usepackage{amsmath, amssymb}
\edef\restoreparindent{\parindent=\the\parindent\relax}
\DeclareSymbolFont{letters}{OML}{ztmcm}{m}{it}
\usepackage{parskip}
\restoreparindent
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\usepackage{verse}

\begin{document}


\poemtitle{True Thomas}

\settowidth{\versewidth}{An example line for measuring the verse width}
\begin{verse}[\versewidth]

    True Thomas lay on Huntlie Bank;\\
    A marvel he did see; \\
    For there he saw a lady bright,\\
    Come riding down by the Eildon tree.\footnote{1}\\
\\

    Her skirt was of the grass-green silk,\\
    Her mantle of the velvet fine; \\
    On every lock of her horse's mane,\\
    Hung fifty silver bells and nine. \\
\\

    True Thomas he pulled off his cap,\\
    And bowed low down on his knee;\\
    \lq\lq All hail, thou mighty Queen of Heaven!\\
    For thy peer on earth could never be.\rq\rq{} \\  \\ 

    \lq\lq O no, O no, Thomas,\rq\rq{} she said, \\ 
    \lq\lq That name does not belong to me; \\ 
    I'm but the Queen of fair Elfland, \\ 
    That hither am come to visit thee. \\  \\ 

    \lq\lq Harp and carp, Thomas,\rq\rq{} she said, \\ 
    \lq\lq Harp and carp along with me; \\ 
    And if ye dare to kiss my lips \\ 
    Sure of your body I will be!\rq\rq{}\\  \\ 

\vin \rq\rq Betide me weal, betide me woe, \\
\vin that weird shall never daunten me!\rq\rq{} \\
\vin Then he has kissed her on the lips, \\
\vin All underneath the Eildon tree. \\ \\ 

    \lq\lq Now ye must go with me,\rq\rq{} she said, \\ 
    \la\la True Thomas, ye must go with me; \\ 
    And ye must serve me seven years, \\ 
    Through weal or woe as may chance to be.\rq\rq{}\\ \\ 

    O they rode on, and farther on, \\ 
    The steed flew swifter than the wind;\\ 
    Until the reached a desert wide, \\ 
    And living land was left behind. \\ \\

    \lq\lq Light down, light down now, Thomas,\rq\rq{} she said, \\ 
    \lq\lq And lean your head upon my knee; \\ 
    Light down, and rest a little space, \\ 
    And I will show you marvels three. \\ \\

    \lq\lq O see ye not yon narrow road, \\ 
    So thick beset with thorns and briers? \\ 
    That is the path of righteousness, \\ 
    Though after it but few enquires. \\ \\

    \lq\lq And see ye not yon broad, broad road, \\ 
    That stretches o\rq{}er the lily leven? \\ 
    That is the path of wickedness, \\ 
    Though some call it the road to heaven. \\ \\ 

    \lq\lq And see ye not yon bonny road, \\ 
    That winds about the green hillside? \\ 
    That is the way to fair Elfland, \\ 
    Where you and I this night must bide. \\ \\ 

    \lq\lq But Thomas, ye shall hold your tongue, \\ 
    Whatever ye may hear or see; \\ 
    For if ye speak word in Elfin land, \\ 
    Ye\rq{}ll ne\rq{}er win back to your own countree!\rq\rq{} \\ \\ 

    O they rode on, and farther on; \\ 
    They waded through rivers above the knee; \\ 
    And they saw neither sun nor moon, \\ 
    But they heard the roaring of a sea. \\ \\ 

    It was mirk, mirk night; there was no star-light;\\ 
    They waded through red blood to the knee; \\ 
    For all the blood that\rq{}s shed on earth, \\
    Runs through the springs o\rq{} that countree. \\ \\ 

    At last they came to a garden green, \\ 
    And she pulled an apple from on high---\\ 
    \lq\lq Take this for thy wages, True Thomas; \\ 
    It will give you thee the tongue that can never lie!\rq\rq{} \\ \\ 

    \lq\lq My tongue is my own, \rq\rq{} True Thomas he said, \\ 
    \lq\lq A goodly gift ye would give to me! \\ 
    I neither could to buy or sell \\ 
    at fair or tryst where I may be. \\  \\ 

    I could neither speak to prince or peer, \\ 
    Nor ask of grace from fair layde.\rq\rq{} \\ 
    \lq\lq Now hold thy peace!\rq\rq{} the lady said, \\ 
    \lq\lq For as I say, so must it be.\rq\rq{} \\ \\ 

    He has gotten a coat of the even cloth, \\ 
    And a pair of shoes of the velvet green; \\ 
    And till seven years were gone and past, \\ 
    True Thomas on earth was never seen.
\end{verse}


\end{document}
